---
title: "Code for Paideia Class, 2023"
author: "Ingrid Zoll"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Installing packages

```{r installing, eval = F}
install.packages("tidyverse")
install.packages("easypackages")

easypackages::install_packages("here", "circlize", "scales", "tvthemes", "remotes", "cowsay",
                 "palmerpenguins", "ggdark", "ggshadow", "fun", "fortunes",
                 "dadjoke", "paletteer", "gggibbous")

remotes::install_github("gadenbuie/ggpomological")
remotes::install_github("sharlagelfand/ggcute")
remotes::install_github("R-CoderDotCom/ggcats@main")
remotes::install_github("R-CoderDotCom/ggdogs@main")
remotes::install_github("R-CoderDotCom/ggbernie@main")
remotes::install_github("rladies/praise")
remotes::install_github("itsrainingdata/emokid")
remotes::install_github("brooke-watson/BRRR")
```

Loading packages

```{r loading, echo=FALSE}
# organization, base packages
library(here)
library(tidyverse)

# data packages
library(palmerpenguins)
library(datasets)

# color/aesthetics packages
library(scales)
library(tvthemes)
library(paletteer)
library(ggpomological)
library(ggcute)
library(ggcats)
library(ggdogs)
library(ggbernie)
library(gggibbous)
library(ggdark)
library(ggshadow)

# graphing packages
library(circlize)

# other packages
library(cowsay)
library(fun)
library(fortunes)
library(dadjoke)
library(praise)
library(emokid)
library(BRRR)
```

## Visually Pleasing Graphs

```{r, echo = F}
set.seed(170513)
n <- 5000
d <- data.frame(a = rnorm(n))
d$b <- .4 * (d$a + rnorm(n))

d$pc <- predict(prcomp(~a+b, d))[,1]

d$density <- fields::interp.surface(
  MASS::kde2d(d$a, d$b), d[,c("a", "b")])

ggplot(d, aes(a, b, color = pc, alpha = 1/density)) +
  geom_point(shape = 16, size = 5, show.legend = FALSE) +
  theme_minimal() +
  scale_color_gradient(low = "#0091ff", high = "#f0650e") +
  scale_alpha(range = c(.05, .25))

set.seed(NULL)
```

From [@drsimonj](https://drsimonj.svbtle.com/pretty-scatter-plots-with-ggplot2)

## Generative Art

```{r, echo = F}
seq(from=-10, to=10, by = 0.05) %>%
expand.grid(x=., y=.) %>%
ggplot(aes(x=(x+pi*sin(y)), y=(y+pi*sin(x)))) +
geom_point(alpha=.1, shape=20, size=1, color="black")+
theme_void()
```

From [@aschinchon](https://twitter.com/aschinchon/status/1614707026415820800)


## Colors

```{r}
gt::info_paletteer(color_pkgs = NULL)
```

## A basic `palmerpenguins` graph

```{r, echo = F}
#library(palmerpenguins)

penguins <- palmerpenguins::penguins

ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g, color = species))+
  geom_point()
```

## Adjust `ggplot2` aesthetics
```{r, echo = F}
ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g, color = species))+
  geom_point()+
  theme_bw()
```

## Adding Other Effects with `ggdark` and `ggshadow`
```{r, echo = F}
#library(ggdark)
#library(ggshadow)

ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g, color = species))+
  geom_glowpoint()+
  dark_theme_linedraw()
```

  
## Different Neon Effect with `ggfx`
```{r, echo = F}
# library(ggfx)

ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g))+
  dark_theme_linedraw()+
  with_outer_glow(geom_point(data=filter(penguins, species == 'Adelie'),fill="white"),
                  colour="dodgerblue", sigma=4, expand=2)+
  with_outer_glow(geom_point(data=filter(penguins, species == 'Gentoo'),fill="white"),
                  colour="deeppink", sigma=4, expand=2)+
  with_outer_glow(geom_point(data=filter(penguins, species == 'Chinstrap'),fill="white"),
                  colour="yellow", sigma=4, expand=2)
```

### Do you ever wish you could make a graph harder to read?
`ggfx`

```{r, echo = F}
# library(ggfx)

pengwing <- ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g, color = species))+
  geom_point()+
  theme_bw()
with_motion_blur(pengwing, sigma =7)
```

  
## Pre-Designed Themes: `ggpomological`
```{r, echo = F}
# library(ggpomological)

ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g, color = species))+
  geom_point()+
  scale_color_pomological()+
  theme_pomological()
```

## Pre-Designed Themes: `ggpomological`
```{r, echo = F}
ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g, color = species))+
  geom_point()+
  scale_color_pomological()+
  theme_pomological("Homemade Apple", 16)
```

## Pre-Designed Themes: `ggcute`
```{r, echo = F}
# library(ggcute)

ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g, color = species))+
  geom_point()+
  scale_color_fairyfloss()+
  theme_fairyfloss()
```

## Pre-Designed Themes: `ggcute`
```{r, echo = F}
ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g, color = species))+
  geom_point()+
  scale_color_sugarpill()+
  theme_sugarpill()
```

## **Game of Thrones Battles**

Chord diagram example with `circlize`

```{r circle, echo = T}
# library(circlize)
```

Game of Thrones battle data from [Kaggle](https://www.kaggle.com/mylesoneill/game-of-thrones?select=battles.csv)

```{r, echo = F}
got <- read_csv(here("Data/got_battles.csv"))

attackers <- got %>%
  select(3, 6:9) %>%
  pivot_longer(cols=c(attacker_1, attacker_2, attacker_3, attacker_4), 
               names_to="val", values_to="attacker") %>%
  filter(!is.na(attacker)) %>%
  select(1, 3)

defenders <- got %>%
  select(3, 10:13) %>%
  pivot_longer(cols=c(defender_1, defender_2, defender_3, defender_4), 
               names_to="val", values_to="defender") %>%
  filter(!is.na(defender)) %>%
  select(1, 3)

battles <- attackers %>%
  full_join(defenders, by="battle_number") %>%
  select(2:3) %>%
  count(attacker, defender) %>%
  rename("from"="attacker", "to"="defender", "value"="n") %>%
  filter(!is.na(to)) 
  
merged <- battles %>%
  pivot_longer(cols=c(from, to), names_to="position", values_to="house", names_repair = "unique") %>%
  count(house) %>%
  arrange(desc(n)) %>%
  filter(n > 2)

top_houses <- merged$house

final_houses <- battles %>%
  filter(from %in% top_houses, to %in% top_houses)
```

```{r}
head(got)
```

### Graph attempt 1 

```{r got1, echo = F}
chordDiagram(battles)
```

### Graph attempt 2

```{r got2, echo = F}
chordDiagram(final_houses)
```

Filtered so both attacking and defending house are involved in 2 or more battles


## Premade Color Palettes
```{r, echo = F}
# library(scales)
```

```{r, echo = T}
# library(tvthemes)
show_col(westeros_pal(palette = "Stark")(6))
```

## Custom Color Palette
```{r, echo = F}
tv_cols <- c("Baratheon" = "#FFD700","Lannister" = "#89080A",
             "Greyjoy" = "#000000","Stark" = "#174d79",
             "Mormont" = "#cadde1","Tully" = "#8b5b45", 
             "Frey" = "#4682b4","Bolton" = "#015202", 
             "Glover" = "#DCDCDC","Karstark" = "#7b906f")
show_col(tv_cols)
```

## Putting it All Together
```{r, echo = F}
chordDiagram(final_houses, grid.col = tv_cols)
```


## Not sure why anyone would need this but it exists
  
```{r, echo = F}
ggplot() + 
  as_reference(
    geom_polygon(aes(c(0, 0, 0), c(0, 0, 0)), colour = NA, fill = 'magenta'), 
    id = "displace_map"
  ) +
  with_displacement(
    geom_text(aes(0.45, 0.5, label = 'word art?'), size = 35, fontface = 'bold'), 
    x_map = ch_red("displace_map"), 
    y_map = ch_blue("displace_map"),
    x_scale = unit(0.025, 'npc'),
    id = "text"
  ) +
  with_blend(
    geom_density_2d_filled(aes(rnorm(1e4, 0.5, 0.2), rnorm(1e4, 0.5, 0.2)), 
                           show.legend = FALSE),
    bg_layer = "text",
    blend_type = "in",
    id = "blended"
  ) +
  scale_fill_viridis_d(option = "plasma")+
  with_shadow("blended", sigma = 3) + 
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), clip = 'off') + 
  labs(x = NULL, y = NULL)
```

## Moon Phase Plots with `gggibbous`

```{r, echo = F}
# library(gggibbous)

moonphase <- subset(lunardist, !is.na(phase))
moonphase$percent <- ifelse(
  moonphase$phase == "new", 0, ifelse(moonphase$phase == "full", 1, 0.5))

moons <- ggplot(lunardist, aes(date, distance)) +
  geom_line(color = "black") +
  # Plotting the lower layer as a full circle also works in most cases
  geom_moon(data = moonphase, ratio = 1, size = 5, fill = "black") +
  geom_moon(
    data = moonphase, aes(ratio = percent),
    size = 5, fill = "yellow", right = moonphase$phase == "first quarter"
  )+
  theme_minimal()

moons
```

## Fun graphing packages: `ggcats`

```{r, echo = F}
# library(ggcats)

ggplot(mtcars) +
  geom_cat(aes(mpg, wt), cat = "pop", size = 3)+
  theme_bw()
```

## Fun graphing packages: `ggdogs`

```{r, echo = F}
# library(ggdogs)

ggplot(mtcars) +
  geom_dog(aes(mpg, wt), dog = "thisisfine", size = 5)+
  theme_bw()
```

## Fun graphing packages: `ggbernie`

```{r, echo = F}
# library(ggbernie)

ggplot(mtcars) +
  geom_bernie(aes(mpg, wt), bernie = "sitting")+
  theme_bw()
```

## Did you know you can animate things?

```{r, echo = F, eval = F}
# library(Ecdat)
data(incomeInequality)

# library(ggcats)
# library(gganimate)

dat <-
  incomeInequality %>%
  select(Year, P99, median) %>%
  rename(income_median = median,
         income_99percent = P99) %>%
  pivot_longer(cols = starts_with("income"),
               names_to = "income",
               names_prefix = "income_")

dat$cat <- rep(NA, 132)

dat$cat[which(dat$income == "median")] <- "nyancat"
dat$cat[which(dat$income == "99percent")] <- rep(c("pop_close", "pop"), 33)

ggplot(dat, aes(x = Year, y = value, group = income, color = income)) +
  geom_line(size = 2) +
  geom_cat(aes(cat = cat), size = 5) +
  xlab("Cats?") +
  ylab("Cats!") +
  theme(legend.position = "none",
        plot.title = element_text(size = 20),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  transition_reveal(Year)+
  theme_bw()
```

## Package: `cowsay`

```{r, echo = F}
# library(cowsay)
```
```{r}
say('time')
```


## Package: `cowsay`
```{r}
say('fortune', 'smallcat')
```

## Package: `cowsay`
```{r}
say('hot diggity', 'frog')
```

## Package: `fun`

- It is possible (though difficult) to play games such as **minesweeper** in R with the package `fun`.

```{r, echo = T, eval = FALSE}
# library(fun)

X11(type = "Xlib")
mine_sweeper(width = 5, height = 5, mines = 5, cheat = FALSE)
```

## Package: `fortunes`

```{r, echo = F}
# library(fortunes)
```

```{r}
fortune()
```

## Package: `praise`

```{r, echo = F}
# library(praise)
```

```{r}
praise()
```


## Dad Jokes
```{r}
# library(dadjoke)
```

```{r}
dadjoke()
```

## Emo Kid
```{r}
# library(emokid)
```

```{r}
iamsad()
```

```{r}
mymood()
```

## Remind me to demonstrate the `BRRR` package later in class...


We can look through more tweets I have saved, I just don't want to deal with adding them
all to the slides 

```{r}
mymood()
```

---
## Sources
